\sysfull

### 2018-08-07 paper/asplos19/background.tex
Final pass on sections 2 and 3 (Carlo)

Energy harvesting devices operate using energy extracted from ambient sources, e.g. radio frequency transmissions, light. These devices elide tethered power and batteries, instead collecting energy into a capacitor, operating when sufficient charge accumulates, and turning off for recharge upon depletion of the buffer. There are several energy harvesting battery-less hardware platforms. For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.

\textbf{Target Hardware of \sys.} \sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}. \sys targets a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM. \sys leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{WAR-affected code execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkpoint()} is a checkpoint of volatile state.}
    \label{fig:war}
\end{figure}

@@

@@

\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or programmer-defined tasks can be both non-terminating and/or inefficient.  If a task (or code between two checkpoints) consumes more than the fixed, maximum energy that the device can buffer, then the task will never be able to complete using buffered energy only.  Such a task is non-terminating, {\em prevents forward progress}, and makes the program deadlock. If the task consumes far less energy than a device can buffer, the system may operate {\em inefficiently}, saving the program state more often than needed. Avoiding excessively costly, non-terminating, tasks and short, high-overhead, tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated. Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse because a large task on one device may be a relatively short task on a device with a larger buffer. \textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which merges multiple tasks to amortize the overhead when tasks are too small, and {\em task downscaling}, which breaks up a task when it is too large to complete.

>>>>

Energy harvesting devices operate using energy extracted from ambient sources, e.g., radio frequency transmissions, light, etc. These devices elide tethered power and batteries, instead collecting energy into a capacitor, operating when sufficient charge accumulates, and turning off for recharge upon depletion of the buffer. There are several energy harvesting battery-less platforms. For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}. 
%
% \textbf{Target Hardware of \sys.}
\sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}. \sys targets a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM. \sys leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.

@@

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{WAR-affected execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, \texttt{checkpoint()} is a checkpoint of volatile state.}
    \label{fig:war}
\end{figure}

@@

\paragraph{Incorrectly Sized Tasks.}
Compiler-inserted checkpoints or programmer-defined tasks can be both non-terminating and/or inefficient.  If a task (or code between two checkpoints) consumes more than the fixed, maximum energy that the device can buffer, then the task will never be able to complete using buffered energy only.  Such a task is non-terminating, {\em prevents forward progress}, and makes the program deadlock. If the task consumes far less energy than a device can buffer, the system may operate {\em inefficiently}, saving the program state more often than needed. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated. Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse because a large task on one device may be a relatively short task on a device with a larger buffer. \textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which merges multiple tasks to amortize the overhead when tasks are too small, and {\em task downscaling}, which breaks up a task when it is too large to complete.

### 2018-08-07 paper/asplos19/systemdescription.tex
Final pass on sections 2 and 3 (Carlo)

\textbf{Programming and Execution Model.}  To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data. Then, he compile the task-based code, and link to \sys's runtime, producing a \sys-enabled binary. The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions. Facilitating efficient task adaptation requires dynamic memory protection which is \sys's \emph{virtual memory manager} through a process we call \emph{dynamic page privatization}.

\textbf{Adaptive Task Scheduler.} \sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task. By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by a task splitting, after repeated failures on a single task, it \emph{avoids the task non-termination problem}. The scheduler uses its recent execution history---i.e., it is hardware independent---as a metric to estimate energy availability and eventually to decide on the coalesced task size. Section~\ref{sec:task_adaptation} describes ATS.

\textbf{Virtual Memory Manager.} \sys can coalesce tasks thanks to its virtual memory manager (VMM). VMM privatizes a group of memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory. VMM keeps all page modifications in non-volatile memory on a coalesced task transition. Section~\ref{sec:memory_virtulaization} outlines VMM.

>>>>

\paragraph{Programming and Execution Model.}
To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data. Then, they compile and link the code against \sys's runtime, producing a \sys-enabled binary. The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions. Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through a process we call \emph{dynamic page privatization}.

\paragraph{Adaptive Task Scheduler.}
\sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task. By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by splitting a task, after it repeatedly failed to complete, it \emph{avoids the task non-termination problem}. The scheduler uses its recent execution history---i.e., it is hardware independent---as a metric to estimate energy availability and eventually to decide on the coalesced task size. Section~\ref{sec:task_adaptation} describes ATS.

\paragraph{Virtual Memory Manager.}
\sys can coalesce tasks thanks to its virtual memory manager (VMM). VMM privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory. VMM keeps all page modifications in non-volatile memory on a coalesced task transition. Section~\ref{sec:memory_virtulaization} outlines VMM.
